Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination

### 2021-01-19 minimax.tex
Some comments

Given a problem size $(m, n)$, a number of questions $k$ and a strategy to test, we first create an “oracle”, representing the true preferences of the agents (randomly generated or coming from real data) and the weights associated with the chair’s scoring rule (randomly generated).
We start with empty knowledge ($\pprofile = (\emptyset, \emptyset, …), W = \W$) about the preference orderings of the agents and the weight differences favored by the chair. We obtain the first question to be asked using the strategy under test. We then use the oracle to answer the question and update the system's knowledge, which is thus used to obtain the next question. This is repeated until $k$ answers have been obtained, computing the resulting MMR values along the way for various values of $k$. We repeat this whole experiment a variable number of times, for a given $(m, n, k)$, and report the average resulting MMR and standard deviation \textit{sd}. The sizes of the considered scenarios are comparable to the ones used by \citet{Cailloux2014}.

@@

\begin{figure}[t]
	\caption{Average MMR in problems of size $(5, 10)$ after $k$ questions.}
	\label{fig:smallSize}
	\begin{tikzpicture}
		\begin{axis}[
			y=8,
			xlabel=Number of Questions,
			ylabel=Avg. Regret,
			ytick={0,2,4,6,8,10},
			xtick distance=10,
			ytick distance=2,
			xtick pos=left,
			ymajorgrids=true,
			ytick style={draw=none},
			ymin=0,
			ymax=11,
			xmin=0,
			xmax=100,
			yticklabels={0,2,4,6,8,10},
			legend style={font=\scriptsize}]
			\addlegendimage{mark=*,teal,mark size=1.5}
			\addlegendimage{mark=triangle*,orange,mark size=1.5}
			\addlegendimage{mark=square*,blue,mark size=1.5}
			\addlegendimage{mark=diamond*,red,mark size=1.5}

@@

%\begin{figure}
%	\caption{Average MMR in problems of size $(5, 10)$ after $k$ questions.}
%	\label{fig:smallSize}
%%	\Description{Graph showing that Pessimistic performs slightly better than Extended pessimistic, itself performing better than Elitist, itself performing much better than Random.}
%	\includegraphics[width=.45\textwidth]{comparison.png}
%\end{figure}
\sisetup{table-number-alignment = center, table-figures-integer=2, table-figures-decimal=1, table-auto-round}
\begin{table}[t]
	\caption{Average MMR in problems of size $(10, 20)$ and geometric weights after $k$ questions.}
	\label{tab:geometricWeights}
	\begin{tabular}{S[table-figures-integer=3, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left, table-figures-integer=1]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left, table-figures-integer=1]}
		\toprule
		{k} & {Pes.} & {sd} & {Eli.} & {sd} \\
		\midrule
		0	&20.0	&0.0	&20.0	&0.0\\
		50	&15.96	&0.54	&17.26	&0.42\\
		100	&12.48	&0.93	&15.6	&0.43\\
		150	&9.58	&1.37	&13.94	&0.76\\
		200	&7.43	&1.25	&10.95	&1.12\\
		250	&5.26	&1.52	&6.6	&0.79\\
		300	&3.47	&1.32	&6.57	&0.79\\
		\bottomrule
	\end{tabular}
\end{table}
\paragraph{Comparison of strategies}
Our first experiment concerns small size situations.
\Cref{fig:smallSize} compares some of our strategies in the case $m = 5, n = 10$ (variations around this size yield similar conclusions), where the results are averaged over $200$ runs.
%(The figure also displays the performance of the Elitist strategy, to which we will come back.)
We see that asking random questions does not allow to reach a low regret level even after having asked 100 questions, whereas a low regret level ($\MMR$ = 1) is reached by Pessimistic before having asked 60 questions.
We also see that Pessimistic performs slightly better than Extended pessimistic, showing that Pessimistic chooses candidate questions wisely; this is good news since Pessimistic is much faster: it takes on average only $16$s for a complete elicitation session (for $m = 5$, $n = 10$ and $100$ questions), while Extended pessimistic takes $50$s. Although their performance is close, Pessimistic performs systematically better in multiple runs of the experiment.
\Cref{sec:appendix} provides a problem of larger size.

@@

\paragraph{Evaluation of Pessimistic Strategy}
\label{sec:lowRegret}
This set of experiments evaluate the Pessimistic strategy in absolute terms. 
We first wonder how many questions should be asked in order to reach a low regret level. We consider a low level to be a tenth of the number of agents: such a regret is equivalent to the difference of score of an alternative $x$ that results from switching from a profile $P$ to a profile $P'$ where ten percent of the agents rank $x$ last instead of first.
The results for various sizes are displayed in \cref{sec:appendix}. 
For the experiments we performed, we see that the number of questions required to reach a low regret level grows approximately linearly with the number of agents and grows less fast than the square of the number of alternatives. Averaging these data by category of value for $m$, and assuming no questions get asked to the chair, we see that in order to reach a low regret level, each agent will have to answer $5.6$ questions for $m = 5$, $17.2$ questions for $m = 10$ and $27.1$ questions for $m = 15$. These numbers have to be contrasted with the estimated number of questions required to elicit the full preference of an agent which are $6.6$ for $m=5$, $24.8$ for $m=10$ and $48.4$ for $m=15$. To obtain these data, we constrained Pessimistic strategy to ask questions to the agents only, and ran it on a single agent problem, with various values for $m$.

@@

%\subsection{Decrease in MMR values}

\begin{figure}[t]
	\caption{Average MMR (normalized by $n$) after $k$ questions with Pessimistic strategy for different data-sets.}
	\label{fig:linearity}
	\begin{tikzpicture}
		\pgfplotsset{
			every axis legend/.append style={
				at={(0.5,1.1)},
				anchor=south
			},
		}
		\begin{axis}[
			y=80,
			legend columns=3,
			xlabel=Number of Questions,
			ylabel=MMR/n,
			ytick={0,0.5,1},
			xtick distance=100,
			xtick pos=left,
			ymajorgrids=true,
			ytick style={draw=none},
			ymin=0,
			ymax=1,
			xmin=0,
			xmax=1000,
			yticklabels={0,0.5,1},
			legend style={font=\scriptsize}]

@@

In \cref{fig:linearity} we can also observe, as expected, that given a problem size the number of questions needed to reach a given MMR is lower when considering a real data-set than a IC generated profile. \commentBN{Add more?}
% do less harm than halving the gain in regret, on average.

>>>>

Given a problem size $(m, n)$, a number of questions $k$ and a strategy to test, we first create an “oracle”, representing the true preferences of the agents (randomly generated or coming from real data) and the weights associated with the chair’s scoring rule (randomly generated).
We start with empty knowledge ($\pprofile = \emptyset, W = \W$) about the preference orderings of the agents and the weight differences favored by the chair. We obtain the first question to be asked using the strategy under test. We then use the oracle to answer the question and update the system's knowledge, which is thus used to obtain the next question. This is repeated until $k$ answers have been obtained, computing the resulting MMR values along the way for various values of $k$. We repeat this whole experiment a variable number of times, for a given $(m, n, k)$, and report the average resulting MMR and standard deviation \textit{sd}. The sizes of the considered scenarios are comparable to the ones used by \citet{Cailloux2014}.

@@

\begin{figure}
	\caption{Average MMR in problems of size $(5, 10)$ after $k$ questions.}
	\label{fig:smallSize}
	\begin{tikzpicture}
		\begin{axis}[
			y=8,
			xlabel=Number of Questions,
			ylabel=Avg. Regret,
			ytick={0,2,4,6,8,10},
			xtick distance=10,
			ytick distance=2,
			xtick pos=left,
			ymajorgrids=true,
			ytick style={draw=none},
			ymin=0,
			ymax=11,
			xmin=0,
			xmax=100,
			yticklabels={0,2,4,6,8,10},
			legend style={font=\scriptsize}]
			\addlegendimage{mark=*,teal,mark size=1.5}
			\addlegendimage{mark=triangle*,orange,mark size=1.5}
			\addlegendimage{mark=square*,blue,mark size=1.5}
			\addlegendimage{mark=diamond*,red,mark size=1.5}

@@

%\begin{figure}
%	\caption{Average MMR in problems of size $(5, 10)$ after $k$ questions.}
%	\label{fig:smallSize}
%%	\Description{Graph showing that Pessimistic performs slightly better than Extended pessimistic, itself performing better than Elitist, itself performing much better than Random.}
%	\includegraphics[width=.45\textwidth]{comparison.png}
%\end{figure}
\sisetup{table-number-alignment = center, table-figures-integer=2, table-figures-decimal=1, table-auto-round}
\begin{table}
	\caption{Average MMR in problems of size $(10, 20)$ and geometric weights after $k$ questions.}
	\label{tab:geometricWeights}
	\begin{tabular}{S[table-figures-integer=3, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left, table-figures-integer=1]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left, table-figures-integer=1]}
		\toprule
		{k} & {Pes.} & {sd} & {Eli.} & {sd} \\
		\midrule
		0	&20.0	&0.0	&20.0	&0.0\\
		50	&15.96	&0.54	&17.26	&0.42\\
		100	&12.48	&0.93	&15.6	&0.43\\
		150	&9.58	&1.37	&13.94	&0.76\\
		200	&7.43	&1.25	&10.95	&1.12\\
		250	&5.26	&1.52	&6.6	&0.79\\
		300	&3.47	&1.32	&6.57	&0.79\\
		\bottomrule
	\end{tabular}
\end{table}
\paragraph{Comparison of strategies}
Our first experiment concerns small size situations.
\Cref{fig:smallSize} compares some of our strategies in the case $m = 5, n = 10$ (variations around this size yield similar conclusions), where the results are averaged over $200$ runs.
%(The figure also displays the performance of the Elitist strategy, to which we will come back.)
We see that asking random questions does not allow to reach a low regret level even after having asked 100 questions, whereas a low regret level ($\MMR$ = 1) is reached by Pessimistic before having asked 60 questions.
We also see that Pessimistic performs slightly better than Extended pessimistic, showing that Pessimistic chooses candidate questions wisely; this is good news since Pessimistic is much faster: it takes on average only $16$s for a complete elicitation session (for $m = 5$, $n = 10$ and $100$ questions), while Extended pessimistic takes $50$s. Although their performance is close, Pessimistic performs systematically better in multiple runs of the experiment.
\Cref{sec:appendix} provides a problem of larger size.

@@

\paragraph{Evaluation of Pessimistic Strategy}
\label{sec:lowRegret}
Our next set of experiments evaluate the Pessimistic strategy in absolute terms. 
We first wonder how many questions should be asked in order to reach a low regret level. We consider a low level to be a tenth of the number of agents: such a regret is equivalent to the difference of score of an alternative $x$ that results from switching from a profile $P$ to a profile $P'$ where ten percent of the agents rank $x$ last instead of first.
The results for various sizes are displayed in \cref{sec:appendix} \commentOC{TODO}. 
For the experiments we performed, we see that the number of questions required to reach a low regret level grows approximately linearly with the number of agents and grows less fast than the square of the number of alternatives. Averaging these data by category of value for $m$, and assuming no questions get asked to the chair, we see that in order to reach a low regret level, each agent will have to answer $5.6$ questions for $m = 5$, $17.2$ questions for $m = 10$ and $27.1$ questions for $m = 15$. These numbers have to be contrasted with the estimated number of questions required to elicit the full preference of an agent which are $6.6$ for $m=5$, $24.8$ for $m=10$ and $48.4$ for $m=15$. To obtain these data, we constrained the Pessimistic strategy to ask questions to the agents only, and ran it on a single agent problem, with various values for $m$.

@@

\begin{figure}
	\caption{Average MMR (normalized by $n$) after $k$ questions with Pessimistic strategy for different data-sets.}
	\label{fig:linearity}
	\begin{tikzpicture}
		\pgfplotsset{
			every axis legend/.append style={
				at={(0.5,1.1)},
				anchor=south
			},
		}
		\begin{axis}[
			y=80,
			legend columns=3,
			xlabel=Number of Questions,
			ylabel=MMR/n,
			ytick={0,0.5,1},
			xtick distance=100,
			xtick pos=left,
			ymajorgrids=true,
			ytick style={draw=none},
			ymin=0,
			ymax=1,
			xmin=0,
			xmax=1000,
			yticklabels={0,0.5,1},
			legend style={font=\scriptsize}]

@@

In \cref{fig:linearity} we can also observe, as expected, that given a problem size the number of questions needed to reach a given MMR is lower when considering a real data-set than a IC generated profile. \commentBN{Add more?} \commentOC{Strange $m=14, n=9$; t-shirts; and comparison between ts and m10n30 is not clearly in favor of real data. I’d suppress the discussion about speed of reduction.}
% do less harm than halving the gain in regret, on average.
